---
title: "Create Data - Disorder"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook creates a data set of disorder calls for the City of St. Louis.

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# tidystl packages
library(stlcsb)

# tidyverse packages
library(dplyr)
library(readr)

# other packages
library(here)
```

## Download Data
We'll start by downloading all csb calls for 2013 through 2017 from the St. Louis Open Data website via the `stlcsb` package:

```{r download-data}
# download
all_calls <- csb_get_data(years = 2013:2017)

# store number of rows
total_rows <- nrow(all_calls)
```

## Subset
Some of these calls were cancelled, so we'll remove them from the data set:

```{r subset}
# subset
all_calls %>% 
  csb_canceled(var = "datecancelled") %>%
  csb_missingXY(varX = srx, varY = sry, newVar = missing) %>%
  filter(missing == FALSE) %>%
  select(-missing) -> all_calls

# store remaining number of rows
remaining_rows <- nrow(all_calls)
```

As an aside, we want to know how many rows were removed by this process:

```{r percent-removed}
remaining_rows/total_rows*100
```

## Cateogrize Data
Next, we'll focus our analysis on particular categories of calls:

```{r categorize-and-subset}
# subset
all_calls %>% 
  csb_categorize(var = problemcode, newVar = category) %>%
  select(requestid, datetimeinit, category, everything()) %>%
  filter(category %in% c("Debris", "Degrade", "Disturbance", "Law")) %>%
  select(datetimeinit, category, probaddress, problemcode, srx, sry) -> focal_calls

# store remaining number of rows
focal_rows <- nrow(focal_calls)
```

As an aside, we want to know how many rows were removed by this process:

```{r percent-remaining}
focal_rows/remaining_rows*100
```

## Write Data to CSV
With these data prepared, we can write them to `.csv`:

```{r write-to-csv}
write_csv(focal_calls, path = here("data", "clean", "disorder.csv"))
```
