---
title: "Create Data - Demographics"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook creates a data set of demographic data for the City of St. Louis at the census tract level.

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# tidyverse packages
library(dplyr)
library(readr)

# other packages
library(here)
library(tidycensus)
```

## Download Census Data
The majority of the demographic variables are available from the U.S. Census Bureau's API, which I'll access using `tidycensus`. 

### Total Population
We'll start with total population per census tract, which comes from table `B01003`.

```{r total-population}
get_acs(geography = "tract", variables = "B01003_001", year = 2017, state = 29, county = 510) %>%
  select(-NAME, -variable) %>%
  rename(
    totalPop = estimate,
    totalPop_moe = moe
  ) -> total_pop
```

### Race
Next, we'll calculate the proportion of each tract's residents that are African American. The underlying race data come from table `B02001`.

```{r}
get_acs(geography = "tract", table = "B02001", year = 2017, state = 29, county = 510, output = "wide") %>%
  select(GEOID, B02001_001E, B02001_001M, B02001_003E, B02001_003M) %>%
  rename(
    raceTotal = B02001_001E,
    raceTotal_moe = B02001_001M,
    black = B02001_003E,
    black_moe = B02001_003M
  ) -> race
```

We'll also preserve the total and black counts, along with their associated margins of error, for the intermediary data set.

### Median Income
Next, we'll get the median household income per census tract, which comes from table `B19013`.

```{r median-income}
get_acs(geography = "tract", variables = "B19013_001", year = 2017, state = 29, county = 510) %>%
  select(-NAME, -variable) %>%
  rename(
    medianInc = estimate,
    medianInc_moe = moe
  ) -> medianInc
```

### Poverty
Next, we'll calculate the proportion of each tract's residents that are living below 100% of the poverty line. The underlying poverty data come from table `B06012`.

```{r poverty}
get_acs(geography = "tract", table = "B06012", year = 2017, state = 29, county = 510, output = "wide") %>%
  select(GEOID, B06012_001E, B06012_001M, B06012_002E, B06012_002M) %>%
  rename(
    pvtyTotal = B06012_001E,
    pvtyTotal_moe = B06012_001M,
    pvty = B06012_002E,
    pvty_moe = B06012_002M
  ) -> poverty
```

As with the race data, the original columns the calculation is based on are preserved here.

### Unemployment
Finally, we'll calculate the proportion of each tract's residents that are unemployed. The underlying employment data come from table `B23025`.

```{r unemployment}
get_acs(geography = "tract", table = "B23025", year = 2017, state = 29, county = 510, output = "wide") %>%
  select(GEOID, B23025_001E, B23025_001M, B23025_007E, B23025_007M) %>%
  rename(
    emplyTotal = B23025_001E,
    emplyTotal_moe = B23025_001M,
    unemply = B23025_007E,
    unemply_moe = B23025_007M
  ) -> employ
```

## Owner Occupied
We'll also calculate the proportion of each tract's residents that are living in owner occupied housing. The underlying residence data come from table `B25026`.

```{r owner-occupied}
get_acs(geography = "tract", table = "B25026", year = 2017, state = 29, county = 510, output = "wide") %>%
  select(GEOID, B25026_001E, B25026_001M, B25026_002E, B25026_002M) %>%
  rename(
    ownTotal = B25026_001E,
    ownTotal_moe = B25026_001M,
    ownOcc = B25026_002E,
    ownOcc_moe = B25026_002M
  ) -> owner
```

### Combine
With our data downloaded, we'll combine the tables:

```{r join}
left_join(total_pop, race, by = "GEOID") %>%
  left_join(., medianInc, by = "GEOID") %>%
  left_join(., owner, by = "GEOID") %>%
  left_join(., poverty, by = "GEOID") %>%
  left_join(., employ, by = "GEOID") -> demos
```

We can now clean-up the global enviornment:

```{r clean-up}
rm(total_pop, race, medianInc, poverty, employ, owner)
```

### Write to CSV
Finally, we'll write the data to a `.csv` file:

```{r write-data}
# write
write_csv(x = demos, path = here("data", "clean", "tract_demographics.csv"))
```
