---
title: "Create Data - Longitudinal Crime Data"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Dependencies
The following packages are needed:

```{r load-dependencies}
# tidyverse
library(dplyr)

# database access
library(DBI)
library(RSQLite)

# other packages
library(here)
```

## Connect to MO_CRIME_Database.sqlite

```{r database-connection}
con <- dbConnect(SQLite(), here("data" , "MO_CRIME_Database", "data", "MO_CRIME_Database.sqlite"))
```

## Construct 
There are a number of database tables that we need. For all of the crimes, we'll need the population for each year. The two `filter()` calls and `select()` execute SQL queries under the hood, and `collect()` brings the data into the global enviornment:

```{r}
# establish connection
pop_db <- tbl(con, "population")

# query and collect
pop_db %>%
  filter(name == "St. Louis") %>%
  filter(year > 1979) %>%
  select(year, count) %>%
  collect() %>%
  rename(pop = count) -> pop

# remove connection
rm(pop_db)
```

Next, we'll query the database again to obtain counts of arson fires. Then we'll join it with the population data and calculate an arson rate:

```{r arson-rate}
# establish connection
arson_db <- tbl(con, "arson")

# query, collect, join, and calculate rate
arson_db %>%
  filter(name == "St. Louis") %>%
  filter(year > 1979) %>%
  select(year, count) %>%
  collect() %>%
  rename(arson = count) %>%
  left_join(pop, ., by = "year") %>%
  mutate(arson_rate = (arson/pop)*100000) -> crime

# remove connection
rm(arson_db, pop)
```

Now, we can continue to add to the `crime` object, using the existing `pop` column to calculate rates.

