---
title: "Create Data - Longitudinal Crime Data"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Dependencies
The following packages are needed:

```{r load-dependencies}
# tidyverse
library(dplyr)
library(readr)

# database access
library(DBI)
library(RSQLite)

# other packages
library(here)
```

## Define Custom Function for SQL Queries
Since all of the database tables are structured identically, we can query them in a uniform fashion. To make the code easier to read, we'll create a function that identifies observations beginning with the given year, collects the records, and returns them as a tibble:

```{r define-query}
query <- function(db_con, firstYear){
  
  # query
  db_con %>%
    filter(name == "St. Louis") %>%
    filter(year >= firstYear) %>%
    select(year, count) %>%
    collect() -> out
  
  # return output
  return(out)
  
}
```

## Connect to MO_CRIME_Database.sqlite

```{r database-connection}
con <- dbConnect(SQLite(), here("data" , "MO_CRIME_Database", "data", "MO_CRIME_Database.sqlite"))
```

## Construct 
### Total Population
There are a number of database tables that we need. For all of the crimes, we'll need the population for each year. The two `filter()` calls and `select()` execute SQL queries under the hood, and `collect()` brings the data into the global enviornment:

```{r}
# establish connection
pop_db <- tbl(con, "population")

# query and collect
pop_db %>%
  query(firstYear = 1979) %>%
  rename(pop = count) -> pop

# remove connection
rm(pop_db)
```

### Arson Rate
Next, we'll query the database again to obtain counts of arson fires. Then we'll join it with the population data and calculate an arson rate:

```{r arson-rate}
# establish connection
arson_db <- tbl(con, "arson")

# query, collect, join, and calculate rate
arson_db %>%
  query(firstYear = 1980) %>%
  rename(arson = count) %>%
  left_join(pop, ., by = "year") %>%
  mutate(arson_rate = (arson/pop)*100000) -> crime

# remove connection
rm(arson_db, pop)
```

Now, we can continue to add to the `crime` object, using the existing `pop` column to calculate rates.

### Violent Crime Rate
Next, we'll calculate a violent crime rate for each year. This involves querying records from four tables, collecting them, and then summing their columns to get a total number of violent crimes before a rate is calculated.

```{r violent-rate}
# establish connections
homicide_db <- tbl(con, "homicide")
agAssault_db <- tbl(con, "agAssault")
rape_db <- tbl(con, "rape")
robbery_db <- tbl(con, "robbery")

# homicide - query and collect
homicide_db %>%
  query(firstYear = 1980) %>%
  rename(homicide = count) -> homicide

# aggrevated assault - query and collect
agAssault_db %>%
  query(firstYear = 1980) %>%
  rename(agAssault = count) -> agAssault

# rape - query and collect
rape_db %>%
  query(firstYear = 1980) %>%
  rename(rape = count) -> rape

# robbery - query and collect
robbery_db %>%
  query(firstYear = 1980) %>%
  rename(robbery = count) -> robbery

# jobin and calculate rates
crime %>%
  left_join(., homicide, by = "year") %>%
  left_join(., agAssault, by = "year") %>%
  left_join(., rape, by = "year") %>%
  left_join(., robbery, by = "year") %>%
  mutate(violent = homicide+agAssault+rape+robbery) %>%
  select(-homicide, -agAssault, -rape, -robbery) %>%
  mutate(violent_rate = (violent/pop)*100000) -> crime

# remove connections
rm(homicide_db, agAssault_db, rape_db, robbery_db,
   homicide, agAssault, rape, robbery)
```

## Larceny Rate
Next, we'll calculate a larceny rate for each year that includes burglaries, larcenies, and motor vehicle thefts. This is the same as the "crimes against property" concept but leaves arsons out, since they are the dependent variable.

```{r larceny-rate}
# establish connections
robbery_db <- tbl(con, "robbery")
larceny_db <- tbl(con, "larceny")
autoTheft_db <- tbl(con, "autoTheft")

# robbery - query and collect
robbery_db %>%
  query(firstYear = 1980) %>%
  rename(robbery = count) -> robbery

# larceny - query and collect
larceny_db %>%
  query(firstYear = 1980) %>%
  rename(larceny = count) -> larceny

# auto theft - query and collect
autoTheft_db %>%
  query(firstYear = 1980) %>%
  rename(autoTheft = count) -> autoTheft

# jobin and calculate rates
crime %>%
  left_join(., robbery, by = "year") %>%
  left_join(., larceny, by = "year") %>%
  left_join(., autoTheft, by = "year") %>%
  mutate(larceny_count = robbery+larceny+autoTheft) %>%
  select(-robbery, -larceny, -autoTheft) %>%
  rename(larceny = larceny_count) %>%
  mutate(larceny_rate = (larceny/pop)*100000) -> crime

# remove connections
rm(robbery_db, larceny_db, autoTheft_db,
   robbery, larceny, autoTheft)
```

## Write Data
Finally, we'll write our crime rate data to `.csv`:

```{r write-crime}
write_csv(crime, path = here("data", "clean", "longitudinal_crime.csv"))
```
