---
title: "Create Data - Longitudinal Demographics"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook calculates relevant demographic metrics for the longitudinal analysis, including population change and percent change in manufacturing jobs at both the MSA and city level.

## Dependencies
The following packages are needed:

```{r load-dependencies}
# tidyverse
library(dplyr)
library(readr)
library(tidyr)

# database access
library(DBI)
library(RSQLite)

# other packages
library(here)
```

## Define Custom Function for SQL Queries
Since all of the database tables are structured identically, we can query them in a uniform fashion. To make the code easier to read, we'll create a function that identifies observations beginning with the given year, collects the records, and returns them as a tibble:

```{r define-query}
query <- function(db_con, firstYear){
  
  # query
  db_con %>%
    filter(name == "St. Louis") %>%
    filter(year >= firstYear) %>%
    select(year, count) %>%
    collect() -> out
  
  # return output
  return(out)
  
}
```

## Connect to MO_CRIME_Database.sqlite
In order to pull total population from the crime database, we need to establish a connection to it:

```{r database-connection}
con <- dbConnect(SQLite(), here("data" , "MO_CRIME_Database", "data", "MO_CRIME_Database.sqlite"))
```

## Construct 
### Total Population
There are a number of database tables that we need. For all of the crimes, we'll need the population for each year. The two `filter()` calls and `select()` execute SQL queries under the hood, and `collect()` brings the data into the global enviornment:

```{r}
# establish connection
pop_db <- tbl(con, "population")

# query and collect
pop_db %>%
  query(firstYear = 1979) %>%
  rename(pop = count) -> pop

# remove connection
rm(pop_db)
```

### Calculate Population Change
In addition to the crime rates, we also want to add a measure of how population numbers changed from year to year (i.e. lagged population).

```{r lagged-pop}
pop %>%
  mutate(pop_delta = pop-lag(pop, n = 1, order_by = year)) %>%
  mutate(pop_delta_pct = pop_delta/lag(pop, n = 1, order_by = year)*100) %>%
  select(year, pop, pop_delta, pop_delta_pct, everything()) -> pop
```

### Manufacturing Jobs
The city job manufacturing data come from the [U.S. Bureau of Economic Analysis](https://www.bea.gov/data/employment/employment-county-metro-and-other-areas). They are in two tables because, of as 2001, the reporting scheme changed. We therefore need to read both in, clean them up, and then join them together. First, we'll wrangle the pre-2001 data:

```{r read-city-pre2001}
# read
city_pre00 <- read_csv(file = here("data", "raw", "employment", "city-1969_to_2000.csv"),
                       skip = 4)

# define variables to reformat
prob_years <- as.character(c(1969:2000))

# subset and convert to long, total
city_pre00 %>%
  filter(Description == "Total employment (number of jobs)" | Description == "Manufacturing") %>%
  mutate(Description = case_when(
    Description == "Total employment (number of jobs)" ~ "total_jobs",
    Description == "Manufacturing" ~ "manu_jobs"
  )) %>%
  select(Description, `1969`:`2000`) %>%
  mutate_at(prob_years, as.numeric) %>%
  pivot_longer(-Description, names_to = "year", values_to = "count") -> city_pre00
```

We'll repeat the exercise with the post-2000 data:

```{r read-city-pre2000}
# read
city_post00 <- read_csv(file = here("data", "raw", "employment", "city-2001_to_2018.csv"),
                       skip = 4)

# define variables to reformat
prob_years <- as.character(c(2001:2018))

# subset and convert to long, total
city_post00 %>%
  filter(Description == "Total employment (number of jobs)" | Description == "Manufacturing") %>%
  mutate(Description = case_when(
    Description == "Total employment (number of jobs)" ~ "total_jobs",
    Description == "Manufacturing" ~ "manu_jobs"
  )) %>%
  select(Description, `2001`:`2018`) %>%
  mutate_at(prob_years, as.numeric) %>%
  pivot_longer(-Description, names_to = "year", values_to = "count") -> city_post00
```

With these wrangled into long data, we can bind them together, and then pivot our numbers into two columns so that this matches the format of the `pop` object (one row per year) and then job them to `pop`:

```{r bind}
bind_rows(city_pre00, city_post00) %>%
  arrange(desc(Description), year) %>%
  pivot_wider(names_from = Description, values_from = count) %>%
  mutate(year = as.numeric(year)) %>%
  full_join(pop, ., by = "year") %>%
  arrange(year) -> demos
```

We can then remove our two intermediary objects:

```{r clean-up}
rm(city_pre00, city_post00, pop)
```

### Manufacturing Jobs, Calculations
Finally, we'll add a ratio of the number of manufacturing jobs to total jobs, and also calculate year to year change:

```{r manufacturing-calculations}
demos %>%
  mutate(manu_delta = manu_jobs-lag(manu_jobs, n = 1, order_by = year)) %>%
  mutate(manu_delta_pct = manu_delta/lag(manu_jobs, n = 1, order_by = year)*100) %>%
  mutate(manu_delta5 = manu_jobs-lag(manu_jobs, n = 5, order_by = year)) %>%
  mutate(manu_delta_pct5 = manu_delta5/lag(manu_jobs, n = 5, order_by = year)*100) -> demos
```

## Write Data
Finally, we'll write our demographic data to `.csv`:

```{r write-crime}
write_csv(demos, path = here("data", "clean", "longitudinal_demos.csv"))
```

