---
title: "Create Data - Longitudinal Analysis Data"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook combines the demographic and crime data to make the analytical data set for the longitudinal data.

## Dependencies
The following packages are needed:

```{r load-dependencies}
# tidyverse
library(dplyr)
library(readr)

# other packages
library(here)
```

## Load Data
There are two tables to be combined:

```{r load-data}
crime <- read_csv(here("data", "clean", "longitudinal_crime.csv"))
demos <- read_csv(here("data", "clean", "longitudinal_demos.csv"))
```

## Subset Years
There are extra rows in the demographic data set leftover from calculating lags. We'll subset those out before joining them:

```{r subset-years}
demos <- filter(demos, year >= 1980)
```

## Join
Next, we'll join the two data sets together:

```{r}
longi <- left_join(crime, demos, by = "year")
```

## Write Data
Finally, we'll write our demographic data to `.csv`:

```{r write-crime}
write_csv(longi, path = here("data", "analysis", "longitudinal_analysis.csv"))
```
